---
title: "HW3_Q4"
format: gfm
editor: visual
---

### Q4 (b) Simulation for one lion

Let $X_n$ denote the lamb position and $Y_n$ denote the lion position.  
The lamb starts at $X_0=0$, the lion starts at $Y_0=10$, and both independently move $+1$ or $-1$ with probability $1/2$ at each time step.

The empirical survival probability is estimated as the proportion of realizations that have not yet been captured by time $t$.

In question (b), we simulate $R=20000$ realizations of the lamb and lion random walks, and estimate the survival probability $S_1(t)$ for $t=0,1,\ldots,10000$.

```{r}
set.seed(123)

R <- 20000
Tmax <- 10000

X <- rep(0, R)      # lamb
Y <- rep(10, R)     # lion

# Logical vector indicating which lambs are still alive
alive <- rep(TRUE, R)

S1 <- numeric(Tmax + 1)
S1[1] <- 1

for (t in 1:Tmax) {

  # Only update realizations that are still alive
  idx <- which(alive)

  if (length(idx) == 0) {
    S1[t + 1] <- 0
    next
  }

  # Vectorized moves for all surviving realizations
  X[idx] <- X[idx] + sample(
    c(-1, 1),
    length(idx),
    replace = TRUE
  )

  Y[idx] <- Y[idx] + sample(
    c(-1, 1),
    length(idx),
    replace = TRUE
  )

  caught <- idx[X[idx] == Y[idx]]
  alive[caught] <- FALSE

  S1[t + 1] <- mean(alive)
}

time <- 0:Tmax
```

fit the slope over $10^2 ≤t≤ 10^4$
```{r}
fit_idx <- time >= 100 &
           time <= 10000 &
           S1 > 0

fit1 <- lm(log(S1[fit_idx]) ~ log(time[fit_idx]))

beta1_hat <- -coef(fit1)[2]

beta1_hat

```

```{r}
erf <- function(x) {
  2 * pnorm(x * sqrt(2)) - 1
}

S1_theory <- rep(NA_real_, length(time))

positive_t <- time > 0

S1_theory[positive_t] <-
  erf(10 / (2 * sqrt(time[positive_t])))

```

```{r}
plot_idx <- time > 0 & S1 > 0

plot(
  time[plot_idx],
  S1[plot_idx],
  log = "xy",
  type = "l",
  lwd = 2,
  col = "black",
  xlab = "Time t",
  ylab = expression(S[1](t)),
  main = sprintf(
    "One lion: fitted beta = %.3f, fit window 10^2 <= t <= 10^4",
    beta1_hat
  )
)

lines(
  time[positive_t],
  S1_theory[positive_t],
  lwd = 2,
  lty = 2,
  col = "blue"
)


fit_time <- time[fit_idx]
fit_survival <- exp(predict(fit1))

lines(
  fit_time,
  fit_survival,
  lwd = 2,
  lty = 3,
  col = "red"
)

legend(
  "bottomleft",
  legend = c(
    "Simulation",
    "Continuum approximation",
    "Power-law fit"
  ),
  col = c("black", "blue", "red"),
  lty = c(1, 2, 3),
  lwd = 2,
  bty = "n"
)
```

## Q4 (c) Simulation for two lion

```{r}
set.seed(123)

R <- 20000
Tmax <- 10000


X  <- rep(0, R)       # lamb
Y1 <- rep(10, R)      # lion 1
Y2 <- rep(10, R)      # lion 2

alive2 <- rep(TRUE, R)

S2 <- numeric(Tmax + 1)
S2[1] <- 1

for (t in 1:Tmax) {

  idx <- which(alive2)

  if (length(idx) == 0) {
    S2[t + 1] <- 0
    next
  }

  # Vectorized random-walk steps
  X[idx] <- X[idx] +
    sample(c(-1, 1), length(idx), replace = TRUE)

  Y1[idx] <- Y1[idx] +
    sample(c(-1, 1), length(idx), replace = TRUE)

  Y2[idx] <- Y2[idx] +
    sample(c(-1, 1), length(idx), replace = TRUE)

  # Captured if the lamb meets either lion
  caught <- idx[
    X[idx] == Y1[idx] |
    X[idx] == Y2[idx]
  ]

  alive2[caught] <- FALSE

  # Proportion still alive
  S2[t + 1] <- mean(alive2)
}

```

```{r}
fit2_idx <- time >= 100 &
            time <= 10000 &
            S2 > 0

fit2 <- lm(
  log(S2[fit2_idx]) ~ log(time[fit2_idx])
)

beta2_hat <- -coef(fit2)[2]

beta2_hat
```

```{r}
S1_sq <- S1^2
```

```{r}
table_times <- c(100, 1000, 10000)

results_table <- data.frame(
  t = table_times,
  S2 = S2[table_times + 1],
  S1_squared = S1[table_times + 1]^2
)

knitr::kable(
  results_table,
  digits = 4,
  col.names = c("t", "S2(t)", "S1(t)^2")
)
```

```{r}
# Continuum prediction from part (b)
erf <- function(x) {
  2 * pnorm(x * sqrt(2)) - 1
}

positive_t <- time > 0

S1_theory <- rep(NA_real_, length(time))
S1_theory[positive_t] <-
  erf(10 / (2 * sqrt(time[positive_t])))

# Valid plotting indices
plot1_idx <- time > 0 & S1 > 0
plot2_idx <- time > 0 & S2 > 0
sq_idx    <- time > 0 & S1_sq > 0

# Start with S1 simulation
plot(
  time[plot1_idx],
  S1[plot1_idx],
  log = "xy",
  type = "l",
  lwd = 2,
  col = "black",
  xlab = "Time t",
  ylab = "Survival probability",
  main = sprintf(
    "beta1 = %.3f, beta2 = %.3f; fit window 10^2 <= t <= 10^4",
    beta1_hat, beta2_hat
  )
)

# Continuum prediction for S1
lines(
  time[positive_t],
  S1_theory[positive_t],
  col = "blue",
  lwd = 2,
  lty = 2
)

# S2 simulation
lines(
  time[plot2_idx],
  S2[plot2_idx],
  col = "red",
  lwd = 2,
  lty = 1
)

# S1(t)^2
lines(
  time[sq_idx],
  S1_sq[sq_idx],
  col = "darkgreen",
  lwd = 2,
  lty = 2
)

# Optional fitted S2 power law
lines(
  time[fit2_idx],
  exp(predict(fit2)),
  col = "purple",
  lwd = 2,
  lty = 3
)

legend(
  "bottomleft",
  legend = c(
    expression(S[1](t)~simulation),
    expression(S[1](t)~continuum),
    expression(S[2](t)~simulation),
    expression(S[1](t)^2),
    expression(S[2](t)~fit)
  ),
  col = c(
    "black",
    "blue",
    "red",
    "darkgreen",
    "purple"
  ),
  lty = c(1, 2, 1, 2, 3),
  lwd = 2,
  bty = "n"
)

```